---
title: "Merge Wordbank imports and export for Redivis"
output:
  html_document:
    toc: true
    df_print: paged
params:
  mode: complete
  skip_processed: true
---

Ingest datasets listed in `datasets.csv` from `raw_data/`, assign IDs consistent
with live `datapages.wordbank`, and write `export/` tables.

**Mode** (`params$mode`):

- **`append`** — pull Redivis, merge only datasets not yet on Redivis, append on upload
- **`complete`** — reimport all manifest raw data, log count/key discrepancies vs Redivis, replace on upload

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE)
library(tidyverse)
library(here)
library(assertthat)

source(here("import", "helpers.R"))
source(here("import", "demographics.R"))
source(here("import", "validate.R"))
source(here("import", "ingest.R"))
source(here("import", "compare.R"))
source(here("import", "ids.R"))

MODE <- params$mode
SKIP_PROCESSED <- isTRUE(params$skip_processed)
stopifnot(MODE %in% c("append", "complete"))
message("Mode: ", MODE, if (SKIP_PROCESSED) " (skip processed)" else "")

RAW <- here("raw_data")
OUT_HARM <- here("harmonized_data")
OUT_EXPORT <- here("export")
dir.create(OUT_HARM, showWarnings = FALSE)
dir.create(OUT_EXPORT, showWarnings = FALSE)
dir.create(file.path(OUT_EXPORT, "item_responses"), showWarnings = FALSE, recursive = TRUE)
```

# Manifest and categories

```{r}
manifest <- read_csv(here("datasets.csv"), show_col_types = FALSE)
categories <- read_csv(
  here("categories.csv"),
  col_names = c("category", "lexical_class", "lexical_category"),
  show_col_types = FALSE
) |>
  distinct()
manifest
```

# Ingest each dataset

One manifest row = one raw triplet. Rows sharing the same
`(dataset_name, dataset_origin_name, language, form)` are combined with
sequential `admin_row` numbering (e.g. Thal longitudinal splits). Directories
like `Dutch_Swingley` hold several distinct datasets — each gets its own group.
Each group is validated against its raw triplets immediately after ingest.
Set `params$skip_processed: true` to load groups that already exist under
`harmonized_data/` instead of re-ingesting raw CSVs.

```{r}
mem.maxVSize(30000)
ingested <- ingest_all_manifest(
  manifest |> arrange(dataset_name),
  RAW,
  categories,
  out_harm = OUT_HARM,
  skip_processed = SKIP_PROCESSED,
  stop_on_fail = F
)
new_parts <- ingested$new_parts
triplet_ranges <- ingested$triplet_ranges
 
walk(names(new_parts), \(nm) {
  message(nm, ": ", nrow(new_parts[[nm]]), " rows")
})
```

Debug example
```{r eval=F}
data_file_path <- "EnglishAmericanWG_Marchman_TL3_data.csv"
rows <- manifest |> filter(data_file == data_file_path)
res <- ingest_dataset_group(rows, RAW, categories, keep_intermediates = TRUE, stop_on_fail = FALSE)
summarize_validation_failures(res$validation)
res$demographic_validation
res$triplets[[data_file_path]]$intermediates$long
```


# Pull existing Redivis core tables

```{r}
existing <- pull_redivis_core()
walk(names(existing), \(nm) message("existing ", nm, ": ", nrow(existing[[nm]])))
```

# Compare ingest to Redivis (pre-merge)

Matches manifest datasets to Redivis using exact keys first, then
`dataset_aliases.csv` for renamed fields (e.g. `German` → `German (German)`,
`Bergman` → `Bergmann`, `CDITwo` → `CDI 2`). Compares administration/child
counts and production/comprehension score distributions. Writes CSVs under
`export/equivalence/`.

```{r}
dataset_aliases <- load_dataset_aliases(here("dataset_aliases.csv"))
equivalence <- compare_manifest_to_redivis(
  existing,
  new_parts,
  dataset_aliases
)
print_dataset_equivalence(equivalence)
write_dataset_equivalence_report(equivalence, file.path(OUT_EXPORT, "equivalence"))
equivalence$comparison
```

# Merge + assign IDs

```{r}
registry <- load_registry(here("id_registry.rds"))
merged <- merge_with_existing(existing, new_parts, registry, mode = MODE)
save_registry(merged$registry, here("id_registry.rds"))
validate_merged(merged$tables, merged$new_item_responses)
if (!is.null(merged$discrepancies)) merged$discrepancies$counts
```

# Export

```{r}
iwalk(merged$tables, \(df, nm) {
  path <- file.path(OUT_EXPORT, paste0(nm, ".csv"))
  write_csv(df, path, na = "")
  message("wrote ", path, " (", nrow(df), " rows)")
})

if (!is.null(merged$upload_deltas)) {
  delta_dir <- file.path(OUT_EXPORT, "deltas")
  dir.create(delta_dir, showWarnings = FALSE)
  iwalk(merged$upload_deltas, \(df, nm) {
    if (nrow(df) == 0) return()
    path <- file.path(delta_dir, paste0(nm, ".csv"))
    write_csv(df, path, na = "")
    message("wrote delta ", path, " (", nrow(df), " rows)")
  })
}

if (nrow(merged$new_item_responses) > 0) {
  merged$new_item_responses |>
    mutate(slug = item_response_slug(language, form)) |>
    group_split(slug) |>
    walk(\(df) {
      slug <- df$slug[[1]]
      path <- file.path(OUT_EXPORT, "item_responses", paste0(slug, ".csv"))
      write_csv(df |> select(-slug), path, na = "")
      message("wrote ", path, " (", nrow(df), " rows)")
    })
}
```

# Optional: upload to Redivis

Requires Redivis authentication. On first use, the `redivis` client opens a
browser OAuth prompt (credentials are cached in `~/.redivis/`).

Upload strategy follows `params$mode`:

- **append** — append delta core-table rows and new `item_responses` files
- **complete** — replace core tables and manifest `item_responses` files

```{r upload, eval=FALSE}
upload_to_redivis(
  merged,
  OUT_EXPORT,
  mode = MODE,
  release = FALSE,
  release_notes = paste("Wordbank import", MODE, Sys.Date())
)
```
